Capstone Movielens Report

### 2025-05-29 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
The helper functions described in the `UMGYDE Model Support Functions` section of the `UMGYD-effect.functions.R` script.

## User+Movie+Genre+Year+Day Effect (UMGYDE) Model
\

### Mathematical Description of the UMGYDE Model
\

@@

::: {.noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, accepting its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

@@

This is a helper function that calculates the _day general effect_ of user ratings on a _given day since the earliest record_ in the `edx` dataset and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)) described below.

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_day_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) function is available in the [UMGYDE Model Support Functions]((https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) script on _GitHub_.
:::

##### Return

A data structure containing information about _Day General Effect_.

@@

This is a helper function that calculates the _day general effect_ of user ratings on a _given day since the earliest record_ in the `edx` dataset using the _5-Fold Cross Validation_ and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv)) described below.

@@

##### Return

A data structure containing information about _Date General Effect_ as result of _5-Fold Cross Validation_.

@@

###### ***de_bias.dat***
\
The _day general effect_ data, which is a data structure returned by any of the following functions described above:

@@

###### ***degree***
\

###### ***span***
\

##### Details
\

The function is a wrapper to the [stats::loess] and calls it under the hood.

@@

##### Return

A data structure containing information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [calc_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) Function {#func.calc_UMGYD_effect}
\

This is a helper function that calculates the _UMGYD Effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGYD_effect](#func.train_UMGYD_effect)) described below.

@@

The function accepts the only parameter `date_general_effect`, which is a data structure returned by any of the following functions described above:

@@

The following is the source code of the function:
```{r eval=FALSE}

@@

::: {.noteblock data-latex=""}
The source code of the [calc_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L69) script on _GitHub_.
:::

##### Return

A data structure containing information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) Function {#func.train_UMGYD_effect}
\

This function is used to train the current model using the _Train Set_ dataset passed in the `train_set` parameter.
The following is the source code of the function:
```{r eval=FALSE}

@@

::: {.noteblock data-latex=""}
The source code of the [train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) script on _GitHub_.
:::

##### Return

A data structure containing information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L84) Function {#func.train_UMGYD_effect.cv}
\

This function is used to train the current model using the _5-Fold Cross Validation_.
The following is the source code of the function:
```{r eval=FALSE}
train_UMGYD_effect.cv <- function(lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_UMGYD_effect.cv
`lambda` is `NA`")
  }

  calc_day_general_effect.cv(lambda) |> calc_UMGYD_effect()  
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L84) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L84) script on _GitHub_.
:::

##### Return

A data structure containing information about _User+Movie+Genre+Year (UMGYD) Effect_.

#### [calc_UMGYD_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L92) Function {#func.calc_UMGYD_effect_MSE}
\

The function calculates the _Mean Squared Error (MSE)_ of the _UMGYDE Model_ for the given _Test Set_ passed in the `test_set` parameter.

@@

  - **test_set:** A _Test Set_ dataset used for _UMGYDE Model_ validation;
  - **UMGYD_effect:** A data structure returned by any of the following functions described above:
    - [calc_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

@@

@@

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_MSE <- function(test_set, UMGYD_effect){
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(UMGYD_effect, by='year') |>
    mutate(resid = rating - clamp(mu + a + b + g + ye)) |> 
    pull(resid) |> mse()
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L92) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L92) script. 
:::

#### [calc_UMGYD_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L102) Function {#func.calc_UMGYD_effect_MSE.cv}
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UMGYD Effect_ data passed in the `UMGYD_effect` parameter.

@@

The function accepts the only parameter: `UMGYD_effect`, which is a data structure returned by the one of the following functions described above:

  - [calc_UMGYD_effect](#func.train_UMGYD_effect);
  - [train_UMGYD_effect](#func.train_UMGYD_effect);
  - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

@@

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_MSE.cv <- function(UMGYD_effect){
  start <- put_start_date()
  UMGYD_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> calc_UMGYD_effect_MSE(UMGYD_effect)
  })
  put_end_date(start)
  put_log1("Function: calc_UMGYD_effect_MSE.cv
Date (Year) Effect MSE values have been computed for the %1-Fold Cross Validation samples.", 
           CVFolds_N)

  mean(UMGYD_effect_MSEs)
}

@@

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L102) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L102) script. 
:::

#### [calc_UMGYD_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L114) Function {#func.calc_UMGYD_effect_RMSE}
\

The function calculates the _Mean Squared Error (MSE)_ of the _UMGYDE Model_ for the given _Test Set_ passed in the `test_set` parameter.

@@

  - **test_set:** A _Test Set_ dataset used for _UMGYDE Model_ validation;
  - **UMGYD_effect:** A data structure returned by any of the following functions described above:
    - [calc_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect](#func.train_UMGYD_effect);
    - [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv).

@@

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGYD_effect_RMSE <- function(test_set, UMGYD_effect){
  mse <- test_set |> calc_UMGYD_effect_MSE(UMGYD_effect)
  sqrt(mse)
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGYD_effect_RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L114) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L114) script. 
:::

#### [calc_UMGYD_effect_RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L118) Function {#func.calc_UMGYD_effect_RMSE.cv}
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UMGYD Effect_ data passed in the `UMGYD_effect` parameter.

>>>>

## User+Movie+Genre+Year+Day Effect (UMGYDE) Model
\
### Mathematical Description of the UMGYDE Model
\

@@

::: {#sec.umgydem.support_functions.note.lambda .noteblock data-latex=""}
Some of the functions described in this section accepts the `lambda` parameter, which we will need later for the _Model Regularization_ method. We use the [mean_reg](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L63) function call as well, which we will also need for the *Regularization* techniques to apply to our models (for details, see the  [mean_reg](#func.mean_reg) function description in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report).
We have already explained that in the [UME Model Regularization] section. 
For now, we omit the `lambda` parameter, implicitly using its default value `lambda = 0`. Let's recall that in this case, the [mean_reg](#func.mean_reg) function is equivalent to the standard R function [base::mean](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean).
:::

@@

This is a helper function that calculates the _Day General Effect_ of user ratings on a _given day since the earliest record_ in the `edx` dataset and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)) described below.

###### ***train_set***
\
The _Train Set_ dataset.

###### ***lambda***
\
As explained [above](#sec.umgydem.support_functions.note.lambda) in this section, this parameter is used for the _Regularization_ techniques. Otherwise, it is omitted, meaning its default (zero) value is used.

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_day_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) function is available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) script on _GitHub_.
:::

##### Return
\
A data frame object representing the _Day General Effect_.

@@

This is a helper function that calculates the _Day General Effect_ of user ratings on a _given day since the earliest record_ in the `edx` dataset using the _5-Fold Cross Validation_ and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv)) described below.

###### ***lambda***
\
As explained [above](#sec.umgydem.support_functions.note.lambda) in this section, this parameter is used for the _Regularization_ techniques. Otherwise, it is omitted, meaning its default (zero) value is used.

@@

##### Return
\
A data frame object representing the _Day General Effect_.

@@

###### ***de_bias.dat***
\
A data frame object representing the _Day General Effect_. It can be returned by the following functions described above:

@@

###### ***degree***
\
A parameter to pass to the internally called [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function (see the [Details](#func.loess_de.details) section below), meaning the degree of the polynomials to be used.

###### ***span***
\
A parameter to pass to the internally called [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function (see the [Details](#func.loess_de.details) section below) to control the degree of smoothing.

::: { #func.loess_de.details .sidebar }
##### Details
\
The function is a wrapper for the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) and calls it under the hood.
:::

@@

##### Return
\
An object of class `loess` returned by the internally called function [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) (see the [Details](#func.loess_de.details) section above).

#### [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) Function {#func.calc_UMGY_SmoothedDay_effect}
\

This is a helper function that calculates the _Smoothed Day Effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)) described below.

@@

::: { #func.calc_UMGY_SmoothedDay_effect.params.day_gen_effect .sidebar }
###### ***day_gen_effect***
\
A data frame object representing the _Day General Effect_. It can be returned by the following functions described above:

  - [calc_day_general_effect](#func.calc_day_general_effect);
  - [calc_day_general_effect.cv](#func.calc_day_general_effect.cv).
:::

###### ***degree***
\
A parameter to pass to the internally called [loess_de](#func.loess_de) function (see the [Details](#func.calc_UMGY_SmoothedDay_effect.details) section below), meaning the degree of the polynomials to be used.

###### ***span***
\
A parameter to pass to the internally called [loess_de](#func.loess_de) function (see the [Details](#func.calc_UMGY_SmoothedDay_effect.details) section below) to control the degree of smoothing.

::: { #func.calc_UMGY_SmoothedDay_effect.details .sidebar }
##### Details
\
Under the hood, the function internally calls the [loess_de](#func.loess_de) described above and constructs the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_ data object based on the object of class `loess` it returns along with the data passed by the [day_gen_effect](#func.calc_UMGY_SmoothedDay_effect.params.day_gen_effect) parameter.
:::

@@

The following is the source code of the function:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect <- function(day_gen_effect, 
                                         degree = NA, 
                                         span = NA){
  put_log2("Function `calc_UMGY_SmoothedDay_effect`: 
Training model using `loess` function with the following parameters:
degree: %1;
span: %2
...",
           degree,
           span)
  # browser()
  fit <- day_gen_effect |> loess_de(degree, span)
  smth_day_effect <- day_gen_effect |> mutate(de_smoothed = fit$fitted)

  put_log2("Function `calc_UMGY_SmoothedDay_effect`: 
Model has been trained using `loess` function with the following parameters:
degree: %1;
span: %2.",
           degree,
           span)

  smth_day_effect
}

@@

::: {.noteblock data-latex=""}
The source code of the [calc_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L74) script on _GitHub_.
:::

##### Return
\
A data frame object representing the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_.

#### [train_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) Function {#func.train_UMGY_SmoothedDay_effect}
\

This is a helper function that calculates the _Smoothed Day Effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)) described below.

###### ***train_set***
\
The _Train Set_ dataset.

###### ***degree***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.details) section below), meaning the degree of the polynomials to be used.

###### ***span***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.details) section below) to control the degree of smoothing.

###### ***lambda***
\
As explained [above](#sec.umgydem.support_functions.note.lambda) in this section, this parameter is used for the _Regularization_ techniques. Otherwise, it is omitted, meaning its default (zero) value is used.

::: { #func.train_UMGY_SmoothedDay_effect.details .sidebar }
##### Details
\
Under the hood, the function internally calls the [calc_day_general_effect](#func.calc_day_general_effect) and then [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) described above, and returns the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_ data object that the latter inner function returns.
:::

The following is the source code of the function:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect <- function(train_set, 
                                          degree = NA, 
                                          span = NA,
                                          lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_UMGY_SmoothedDay_effect
`lambda` is `NA`")
  }

  train_set |> 
    calc_day_general_effect(lambda) |>
    train_UMGY_SmoothedDay_effect(degree, span)
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_SmoothedDay_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) script on _GitHub_.
:::

##### Return
\
A data frame object representing the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_.

#### [train_UMGY_SmoothedDay_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L111) Function {#func.train_UMGY_SmoothedDay_effect.cv}
\

This is a helper function that calculates the _Smoothed Day Effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv)) described below.

###### ***degree***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.cv.details) section below), meaning the degree of the polynomials to be used.

###### ***span***
\
A parameter to pass to the internally called [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.cv.details) section below) to control the degree of smoothing.

###### ***lambda***
\
As explained [above](#sec.umgydem.support_functions.note.lambda) in this section, this parameter is used for the _Regularization_ techniques. Otherwise, it is omitted, meaning its default (zero) value is used.

::: { #func.train_UMGY_SmoothedDay_effect.cv.details .sidebar }
##### Details
\
Under the hood, the function internally calls the [calc_day_general_effect.cv](#func.calc_day_general_effect.cv) and then [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) described above, and returns the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_ data object that the latter inner function returns.
:::

The following is the source code of the function:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect.cv <- function(degree = NA, 
                                             span = NA,
                                             lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_UMGY_SmoothedDay_effect.cv
`lambda` is `NA`")
  }

  calc_day_general_effect.cv.cv(lambda) |>
    calc_UMGY_SmoothedDay_effect(degree, span)
}

@@

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_SmoothedDay_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L111) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L111) script on _GitHub_.
:::

##### Return
\
A data frame object representing the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_.

#### [UMGY_SmoothedDay_effect.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L123) Function {#func.UMGY_SmoothedDay_effect.predict}
\

This is a helper function that predicts user rating values on the _Test Set_ passed by the [test_set](#func.UMGY_SmoothedDay_effect.predict.params.test_set) parameter, taking into account the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_ passed by the [day_smoothed_effect](#func.UMGY_SmoothedDay_effect.predict.params.day_smoothed_effect) parameter.

@@

::: { #func.UMGY_SmoothedDay_effect.predict.params.test_set .sidebar }
###### ***test_set***
\
A _Test Set_ dataset for the _UMGYDE Model_ validation.
:::

::: { #func.UMGY_SmoothedDay_effect.predict.params.day_smoothed_effect .sidebar }

###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

  - [calc_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

:::

Below is the source code of the function:
```{r eval=FALSE}
UMGY_SmoothedDay_effect.predict <- function(test_set, day_smoothed_effect) {
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(rglr.UMGY_effect, by='year') |>
    left_join(day_smoothed_effect, by='days') |>
    mutate(predicted = clamp(mu + a + b + g + 
                               ifelse(is.na(ye), 0, ye) +
                               ifelse(is.na(de_smoothed), 
                                      0, 
                                      de_smoothed))) |> 
    select(userId, movieId, timestamp, rating, predicted)
}
```

::: {.noteblock data-latex=""}
The source code of the [UMGY_SmoothedDay_effect.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L123) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L123) script on _GitHub_.
:::

##### Return
\

A data frame object containing predicted values.

@@

#### [calc_UMGY_SmoothedDay_effect.MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L139) Function {#func.calc_UMGY_SmoothedDay_effect.MSE}
\

The function calculates the _Mean Squared Error (MSE)_ of the _UMGYDE Model_ for the given _Test Set_ passed by the [test_set](#func.calc_UMGY_SmoothedDay_effect.MSE.params.test_set) parameter.

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.params.test_set .sidebar }
###### ***test_set***
\
A _Test Set_ dataset for the _UMGYDE Model_ validation.
:::

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.params.day_smoothed_effect .sidebar }

###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

  - [calc_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

:::

@@

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.MSE <- function(test_set, day_smoothed_effect) {
  test_set |>
    UMGY_SmoothedDay_effect.predict(day_smoothed_effect) |>
    mutate(resid = rating - predicted) |> 
    pull(resid) |> mse()
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGY_SmoothedDay_effect.MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L139) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L139) script on _GitHub_. 
:::

##### Return
\
The _Mean Squared Error (MSE)_ value of the _UMGYDE Model_ computed on the _Test Set_ passed by the [test_set](#func.calc_UMGY_SmoothedDay_effect.MSE.params.test_set) parameter.

#### [calc_UMGY_SmoothedDay_effect.MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L145) Function {#func.calc_UMGY_SmoothedDay_effect.MSE.cv}
\

The function calculates the _5-Fold Cross Validation MSE_ result for the _UMGYD Effect_ data passed by the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.cv.params.day_smoothed_effect) parameter.

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.cv.params.day_smoothed_effect .sidebar }

###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

  - [calc_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

:::

@@

Below is a simplified version of the function's source code:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.MSE.cv <- function(day_smoothed_effect){
  smth_day_effect_MSEs <- sapply(edx_CV, function(cv_fold_dat){
    cv_fold_dat$validation_set |> 
      calc_UMGY_SmoothedDay_effect.MSE(day_smoothed_effect)
  })

  mean(smth_day_effect_MSEs)
}

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_UMGY_SmoothedDay_effect.MSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L145) function is available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L145) script on _GitHub_.
:::

##### Return
\
The _Mean Squared Error (MSE)_ value of the _UMGYDE Model_ computed as the mean of the _5-Fold Cross Validation_ results.

#### [calc_UMGY_SmoothedDay_effect.RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L161) Function {#func.calc_UMGY_SmoothedDay_effect.RMSE}
\

The function calculates the _Root Mean Squared Error (RMSE)_ of the _UMGYDE Model_ for the given _Test Set_ passed by the [test_set](#func.calc_UMGY_SmoothedDay_effect.RMSE.params.test_set) parameter.

@@

::: { #func.calc_UMGY_SmoothedDay_effect.RMSE.params.test_set .sidebar }
###### ***test_set***
\
A _Test Set_ dataset for the _UMGYDE Model_ validation.
:::

::: { #func.calc_UMGY_SmoothedDay_effect.RMSE.params.day_smoothed_effect .sidebar }
###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

  - [calc_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect);
  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

:::

@@

The source code of the function is provided below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.RMSE <- function(test_set, day_smoothed_effect){
  day_smth_effect_MSe <- test_set |> 
    calc_UMGY_SmoothedDay_effect.MSE(day_smoothed_effect)

  sqrt(day_smth_effect_MSe)
}
```

::: {.noteblock data-latex=""}
The source code of the function [calc_UMGY_SmoothedDay_effect.RMSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L161) is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L161) script. 
:::

##### Return
\
The _Root Mean Squared Error (RMSE)_ value of the _UMGYDE Model_ computed on the _Test Set_ passed by the [test_set](#func.calc_UMGY_SmoothedDay_effect.RMSE.params.test_set) parameter.

#### [calc_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L167) Function {#func.calc_UMGY_SmoothedDay_effect.RMSE.cv}
\

The function calculates the _5-Fold Cross Validation RMSE_ result for the _UMGYD Effect_ data passed by the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv.params.day_smoothed_effect) parameter.
